---
title: "Multiple Linear Regression: A Macro-Augmented Factor Model of NVIDIA's Daily Returns"
author: "Jan Zika"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
    theme: cosmo
  pdf:
    toc: true
    number-sections: true
    fig-width: 7
    fig-height: 4.5
execute:
  warning: false
  message: false
  echo: true
---

```{r setup}
#| include: false
# Source the data pipeline (fetched from the FRED + market-price APIs) and load
# the processed panels it writes to data/cache/processed/.
repo <- normalizePath(file.path(getwd(), "..", ".."), mustWork = FALSE)
if (!file.exists(file.path(repo, "config", "symbols.yml"))) repo <- getwd()
knitr::opts_knit$set(root.dir = repo)

suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(readr); library(ggplot2)
  library(broom); library(car); library(lmtest); library(knitr)
})

source(file.path(repo, "R", "config.R"))

prices_wide <- readr::read_csv(file.path(processed_dir(), "prices_wide.csv"),
                               show_col_types = FALSE)
macro_wide  <- readr::read_csv(file.path(processed_dir(), "macro_wide.csv"),
                               show_col_types = FALSE)
```

```{r build-model}
#| include: false
# --- Daily simple returns from the adjusted-close panel ---
rets <- prices_wide |> arrange(date)
ret_cols <- setdiff(names(rets), "date")
for (cc in ret_cols) rets[[cc]] <- rets[[cc]] / dplyr::lag(rets[[cc]]) - 1
rets <- rets[-1, ]

# --- Macro factors from the FRED panel ---
# DTB3 -> daily risk-free rate; spreads and VIX entered as daily first
# differences so each predictor is a stationary daily change, matched to returns.
macro <- macro_wide |> arrange(date) |>
  mutate(
    rf_daily = (DTB3 / 100) / 252,
    d_term   = T10Y3M - dplyr::lag(T10Y3M),
    d_credit = BAA10Y - dplyr::lag(BAA10Y),
    d_vix    = VIXCLS - dplyr::lag(VIXCLS)
  ) |>
  select(date, rf_daily, d_term, d_credit, d_vix)

# --- Assemble the model frame: one row per trading day ---
model_df <- rets |>
  inner_join(macro, by = "date") |>
  mutate(
    NVDA_excess  = NVDA - rf_daily,   # outcome: NVIDIA daily excess return
    Market       = VTI - rf_daily,    # market excess return (the CAPM term)
    Size         = IWM - SPY,         # small-cap minus large-cap
    Value        = IWD - IWF,         # value minus growth
    Sector       = XSD - VTI,         # semiconductor tilt over the market
    TermSpread   = d_term,            # daily change in the 10Y-3M yield slope
    CreditSpread = d_credit,          # daily change in the Baa-10Y credit spread
    Volatility   = d_vix              # daily change in the VIX
  ) |>
  select(date, NVDA_excess, Market, Size, Value, Sector,
         TermSpread, CreditSpread, Volatility) |>
  tidyr::drop_na()

model <- lm(NVDA_excess ~ Market + Size + Value + Sector +
              TermSpread + CreditSpread + Volatility, data = model_df)

s   <- summary(model)
co  <- coef(model)
n_obs <- nrow(model_df)
date_min <- format(min(model_df$date)); date_max <- format(max(model_df$date))
fmt <- function(x, d = 4) formatC(x, format = "f", digits = d)
```

## Research Question

**Which broad market, style, sector, and macroeconomic forces move NVIDIA's daily
stock return, and how much of that return do they jointly explain?**

NVIDIA (NVDA) is one of the most heavily traded stocks in the market, and its
day-to-day moves are widely treated as a barometer for the technology and
semiconductor sectors. This analysis asks how much of NVIDIA's daily return is
*systematic* — driven by forces that move the whole market — versus
*idiosyncratic* — specific to the company itself. The answer matters to anyone
holding or hedging the stock: it tells them what NVIDIA is really exposed to, and
how much of its risk could in principle be diversified or hedged away.

The outcome is NVIDIA's daily **excess return** (its return above the risk-free
rate). The seven predictors span the market itself, two investment *style*
factors (size and value), a *sector* factor, and three *macroeconomic* signals
(the yield-curve slope, the corporate credit spread, and market-wide volatility).

## Dataset and Variables

**Source.** All data is pulled programmatically through public market-data APIs.
Daily adjusted closing prices come from a market-price API; the macroeconomic and
interest-rate series come from the U.S. Federal Reserve's FRED API. A small R
pipeline (`R/update_data.R`) fetches every series, caches it, and assembles the
clean daily panel used here.

**What each row represents.** One **trading day**. Each row pairs NVIDIA's return
for that day with the value of every market, style, sector, and macro factor on
the same day.

**Number of observations.** The model is estimated on **`r n_obs` trading days**,
spanning **`r date_min`** to **`r date_max`**.

**Variables.**

| Role | Variable | What it represents |
|------|----------|--------------------|
| Outcome | **NVDA excess return** | NVIDIA's daily return minus the risk-free rate. |
| Predictor 1 | **Market** | The total stock market's excess return (a broad-market ETF minus the risk-free rate). This is the classic CAPM market factor. |
| Predictor 2 | **Size** | Small-cap return minus large-cap return. Positive when smaller companies outperform. |
| Predictor 3 | **Value** | Value-stock return minus growth-stock return. Positive when cheap "value" stocks beat fast-growing "growth" stocks. |
| Predictor 4 | **Sector** | The semiconductor sector's return above the market. Captures chip-industry-specific moves. |
| Predictor 5 | **TermSpread** | The daily change in the 10-year-minus-3-month Treasury yield slope. A rising slope often signals improving growth expectations. |
| Predictor 6 | **CreditSpread** | The daily change in the Baa-corporate-minus-10-year-Treasury yield spread. A rising spread signals growing financial stress. |
| Predictor 7 | **Volatility** | The daily change in the VIX, the market's "fear gauge." |

The risk-free rate is the 3-month Treasury bill rate. The spreads and the VIX
enter as **daily changes** rather than levels, so each predictor is a stationary
daily signal that lines up naturally with a daily return.

## The Regression Model Equation

The general form of a multiple linear regression with seven predictors is:

$$
Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + \beta_4 X_4
      + \beta_5 X_5 + \beta_6 X_6 + \beta_7 X_7 + \varepsilon
$$

where $Y$ is the outcome, $\beta_0$ is the intercept, each $\beta_k$ is the
coefficient on predictor $X_k$, and $\varepsilon$ is the error term. Written with
our own variable names:

$$
\text{NVDA excess} = \beta_0 + \beta_1\,\text{Market} + \beta_2\,\text{Size}
  + \beta_3\,\text{Value} + \beta_4\,\text{Sector}
  + \beta_5\,\text{TermSpread} + \beta_6\,\text{CreditSpread}
  + \beta_7\,\text{Volatility} + \varepsilon
$$

After fitting the model, the estimated equation (using the actual coefficients
computed below) is:

$$
\widehat{\text{NVDA excess}} = `r fmt(co[1])`
  + `r fmt(co[2])`\,\text{Market}
  `r fmt(co[3])`\,\text{Size}
  `r fmt(co[4])`\,\text{Value}
  + `r fmt(co[5])`\,\text{Sector}
  + `r fmt(co[6])`\,\text{TermSpread}
  `r fmt(co[7])`\,\text{CreditSpread}
  + `r fmt(co[8])`\,\text{Volatility}
$$

## Running the Model in R

```{r run-model}
#| code-fold: false
model <- lm(NVDA_excess ~ Market + Size + Value + Sector +
              TermSpread + CreditSpread + Volatility, data = model_df)
summary(model)
```

```{r coef-table}
broom::tidy(model) |>
  mutate(across(where(is.numeric), \(x) round(x, 5))) |>
  kable(caption = "Estimated coefficients, standard errors, t-statistics, and p-values.")
```

## Interpreting the Model Summary

### A. Coefficients

Each coefficient is the average change in NVIDIA's daily excess return associated
with a one-unit increase in that predictor, **holding the other six predictors
constant**.

- **Market (`r fmt(co["Market"])`).** Positive. When the market's excess return
  rises by one unit, NVIDIA's expected excess return rises by about
  `r fmt(co["Market"])`. A value above 1 means NVIDIA is a **high-beta** stock: it
  amplifies market moves, climbing more than the market in up days and falling
  more in down days.
- **Size (`r fmt(co["Size"])`).** Negative. When small-cap stocks beat large-caps,
  NVIDIA tends to lag slightly — consistent with NVIDIA trading like a large-cap.
- **Value (`r fmt(co["Value"])`).** Strongly negative. When value stocks beat
  growth stocks, NVIDIA falls. This is the model's clearest style signal: NVIDIA
  behaves like a **growth** stock, moving opposite to the value-minus-growth factor.
- **Sector (`r fmt(co["Sector"])`).** Positive. Independent of the broad market,
  when semiconductors outperform, NVIDIA rises with its sector — exactly what we
  expect from a chip maker.
- **TermSpread (`r fmt(co["TermSpread"])`).** Small. A one-day steepening of the
  yield curve has little measured association with NVIDIA's return once the other
  factors are held constant.
- **CreditSpread (`r fmt(co["CreditSpread"])`).** Negative. On days when the
  corporate credit spread widens — a sign of rising financial stress — NVIDIA
  tends to fall, even after accounting for the market.
- **Volatility (`r fmt(co["Volatility"])`).** Very small once the market factor is
  in the model, because changes in the VIX and market returns carry overlapping
  information.

### B. P-values

```{r pvals}
broom::tidy(model) |>
  transmute(term, estimate = round(estimate, 5), p.value = signif(p.value, 3),
            significant_5pct = ifelse(p.value < 0.05, "yes", "no")) |>
  kable(caption = "Which predictors show statistically significant relationships.")
```

A small p-value is evidence that a predictor's true coefficient is different from
zero. The **market, size, value, and sector** factors are all highly significant
(p-values far below 0.05): the data give strong evidence that each is genuinely
related to NVIDIA's return. The **credit spread** is significant at the 5% level,
showing measurable sensitivity to financial-stress days. By contrast, the **term
spread** and the **change in the VIX** are *not* statistically significant: once
the market and sector factors are present, these two macro signals add no
detectable independent relationship with NVIDIA's daily return. A non-significant
predictor does not prove there is no relationship — only that this data set shows
no clear evidence of one.

### C. R-squared

The R-squared value is **`r fmt(s$r.squared, 3)`**, which means the model explains
approximately **`r round(100 * s$r.squared)`%** of the day-to-day variation in
NVIDIA's excess return. In other words, roughly that share of NVIDIA's daily moves
is *systematic* — traceable to broad market, style, sector, and macro forces —
while the remaining **`r round(100 * (1 - s$r.squared))`%** is **idiosyncratic**:
specific to NVIDIA (company news, earnings, product announcements) and not
captured by these factors.

### D. Adjusted R-squared

The adjusted R-squared is **`r fmt(s$adj.r.squared, 3)`**. Adjusted R-squared
penalizes the model for each extra predictor, so it only rises when a new variable
earns its place. Here it is almost identical to the ordinary R-squared
(`r fmt(s$r.squared, 3)`), telling us the seven predictors together are not
padding the model with noise — the explanatory power is real and not an artifact
of simply adding variables. The fact that the two values barely differ is a sign
of a healthy, well-specified model.

## Residual Errors and Model Fit

A **residual** is the difference between the actual value and the model's
prediction:

$$\text{Residual} = \text{Actual} - \text{Predicted}$$

A residual tells us how far off the model was on a single day. A **positive**
residual means the model **underpredicted** (NVIDIA did better than expected); a
**negative** residual means it **overpredicted** (NVIDIA did worse than expected).
For example, if NVIDIA's actual excess return on a day was 3% and the model
predicted 2%, the residual is +1%.

## Residual Plot

```{r resid-plot}
#| fig-cap: "Residuals versus fitted values. A random horizontal band around zero indicates a well-behaved linear model."
model_df$predicted <- fitted(model)
model_df$residuals <- residuals(model)

ggplot(model_df, aes(x = predicted, y = residuals)) +
  geom_point(alpha = 0.25, color = "#2c7fb8") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(se = FALSE, color = "#d95f02", linewidth = 0.7) +
  labs(title = "Residuals vs. Predicted Values (NVIDIA excess return)",
       x = "Predicted excess return", y = "Residual") +
  theme_minimal()
```

```{r diagnostic-plots}
#| fig-cap: "Standard regression diagnostics: residuals vs fitted, Normal Q-Q, scale-location, and residuals vs leverage."
par(mfrow = c(2, 2)); plot(model); par(mfrow = c(1, 1))
```

The residuals are scattered fairly evenly above and below zero with no obvious
trend across the range of predictions, so the orange smoother stays close to the
dashed zero line — the **linearity** assumption looks reasonable. The visible
feature is that the *spread* of the residuals widens on high-volatility days:
large surprises cluster together, a familiar property of stock returns. The model
does not systematically over- or under-predict in any region; its errors are
roughly centered on zero everywhere, but it is least precise during turbulent
market periods. The Q-Q plot shows heavier tails than a normal distribution —
extreme days (large jumps) happen more often than a bell curve would predict.

## Model Assumptions and Limitations

```{r diagnostics}
vif_vals <- car::vif(model)
bp <- lmtest::bptest(model)
dw_lag1 <- cor(head(residuals(model), -1), tail(residuals(model), -1))
```

- **Linearity.** The residual plot shows no strong curvature, so a straight-line
  relationship between the factors and NVIDIA's return is a defensible
  approximation.
- **Residual randomness / independence.** The day-to-day autocorrelation of the
  residuals is only about `r fmt(dw_lag1, 3)`, essentially zero, so the model is
  not leaving obvious predictable structure in its errors.
- **Outliers.** Extreme trading days (earnings surprises, market shocks) appear in
  the tails of the Q-Q plot. They do not dominate the fit given the large sample,
  but they are the days the model predicts least well.
- **Multicollinearity.** All variance-inflation factors are low (the largest is
  about `r fmt(max(vif_vals), 1)`, well under the usual threshold of 5), so the
  predictors are not redundant and each coefficient is estimated on its own
  distinct information.
- **Constant error variance.** A Breusch–Pagan test rejects constant variance
  (p `r ifelse(bp$p.value < 0.001, "< 0.001", paste("=", fmt(bp$p.value, 3)))`):
  the model's errors are larger on volatile days. This **heteroskedasticity** does
  not bias the coefficients, but it means the reported standard errors slightly
  overstate the precision of the estimates. The large significances here are wide
  enough that this caveat does not change which factors matter.
- **Causation.** This model shows **association, not causation**. A significant
  coefficient means a factor *moves together with* NVIDIA's return, not that it
  *causes* it. These factors and NVIDIA are all driven by the same underlying
  market forces in real time.

## The "So What?"

This model turns a noisy stream of daily NVIDIA returns into a clear story.
About **`r round(100 * s$r.squared)`%** of NVIDIA's daily movement is **systematic**
— explained by where the broad market, growth-versus-value sentiment, and the
semiconductor sector went that day. The single most important driver is the
**market itself**: with a beta of `r fmt(co["Market"], 2)`, NVIDIA amplifies
market moves, so it is a high-risk, high-reward holding rather than a defensive
one. Its strong **growth** tilt (the large negative Value coefficient) and its
clear **semiconductor** loading confirm what investors intuit — NVIDIA lives or
dies with growth-stock sentiment and the chip cycle. The significant **credit
spread** effect adds that NVIDIA also feels broad financial stress.

The practical takeaways: an investor who already owns the market and the
semiconductor sector is, to a large degree, *already* holding NVIDIA's risk —
which is useful for hedging and diversification decisions. But the remaining
**`r round(100 * (1 - s$r.squared))`%** of NVIDIA's daily variation is
**idiosyncratic** and cannot be explained by these broad factors; it is the part
driven by company-specific news. Anyone using this model should remember three
things: it explains *association*, not cause; it is least reliable precisely on
the most turbulent days; and it describes the average relationship over this
sample, which can shift as market regimes change. It is a strong tool for
*understanding* what NVIDIA is exposed to — not a crystal ball for predicting
tomorrow's price.
