---
title: "Multiple Linear Regression: A Macro-Augmented Factor Model of Micron's Daily Returns"
author: "Jan Zika"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
    theme: cosmo
  pdf:
    toc: true
    number-sections: true
    fig-width: 7
    fig-height: 4.5
    include-in-header:
      text: |
        \usepackage{fvextra}
        \fvset{breaklines,breaknonspaceingroup,breaksymbolleft={}}
        \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,breaknonspaceingroup,breaksymbolleft={},commandchars=\\\{\}}
execute:
  warning: false
  message: false
  echo: true
---

```{r setup}
#| include: false
# Source the data pipeline (fetched from the FRED + market-price APIs) and load
# the processed panels it writes to data/cache/processed/.
repo <- normalizePath(file.path(getwd(), "..", ".."), mustWork = FALSE)
if (!file.exists(file.path(repo, "config", "symbols.yml"))) repo <- getwd()
knitr::opts_knit$set(root.dir = repo)

suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(readr); library(ggplot2)
  library(broom); library(car); library(lmtest); library(knitr)
  library(strucchange); library(sandwich); library(zoo)
})

source(file.path(repo, "R", "config.R"))

prices_wide <- readr::read_csv(file.path(processed_dir(), "prices_wide.csv"),
                               show_col_types = FALSE)
macro_wide  <- readr::read_csv(file.path(processed_dir(), "macro_wide.csv"),
                               show_col_types = FALSE)
```

```{r build-model}
#| include: false
# --- Daily simple returns from the adjusted-close panel ---
rets <- prices_wide |> arrange(date)
ret_cols <- setdiff(names(rets), "date")
for (cc in ret_cols) rets[[cc]] <- rets[[cc]] / dplyr::lag(rets[[cc]]) - 1
rets <- rets[-1, ]

# --- Macro factors from the FRED panel ---
# DTB3 -> daily risk-free rate; spreads and VIX entered as daily first
# differences so each predictor is a stationary daily change, matched to returns.
macro <- macro_wide |> arrange(date) |>
  mutate(
    rf_daily = (DTB3 / 100) / 252,
    d_term   = T10Y3M - dplyr::lag(T10Y3M),
    d_credit = BAA10Y - dplyr::lag(BAA10Y),
    d_vix    = VIXCLS - dplyr::lag(VIXCLS)
  ) |>
  select(date, rf_daily, d_term, d_credit, d_vix)

# --- Assemble the model frame: one row per trading day ---
model_df <- rets |>
  inner_join(macro, by = "date") |>
  mutate(
    MU_excess    = MU - rf_daily,     # outcome: Micron daily excess return
    Market       = VTI - rf_daily,    # market excess return (the CAPM term)
    Size         = IWM - SPY,         # small-cap minus large-cap
    Value        = IWD - IWF,         # value minus growth
    Sector       = XSD - VTI,         # semiconductor tilt over the market
    TermSpread   = d_term,            # daily change in the 10Y-3M yield slope
    CreditSpread = d_credit,          # daily change in the Baa-10Y credit spread
    Volatility   = d_vix,             # daily change in the VIX
    NVDA_ret     = NVDA               # NVIDIA daily return (AI-demand bellwether)
  ) |>
  select(date, MU_excess, Market, Size, Value, Sector,
         TermSpread, CreditSpread, Volatility, NVDA_ret) |>
  tidyr::drop_na()

model <- lm(MU_excess ~ Market + Size + Value + Sector +
              TermSpread + CreditSpread + Volatility, data = model_df)

s   <- summary(model)
co  <- coef(model)
n_obs <- nrow(model_df)
date_min <- format(min(model_df$date)); date_max <- format(max(model_df$date))
fmt <- function(x, d = 4) formatC(x, format = "f", digits = d)
```

## Research Question

**Which broad market, style, sector, and macroeconomic forces move Micron's daily
stock return, and how much of that return do they jointly explain?**

Micron Technology (MU) is the leading U.S. maker of memory chips — the DRAM and
NAND that go into everything from data-center accelerators to phones — and its
moves are watched closely as a read on the memory cycle and on demand for
AI-driven computing. This analysis asks how much of Micron's daily return is
*systematic* — driven by forces that move the whole market — versus
*idiosyncratic* — specific to the company itself. The answer matters to anyone
holding or hedging the stock: it tells them what Micron is really exposed to, and
how much of its risk could in principle be diversified or hedged away.

The outcome is Micron's daily **excess return** (its return above the risk-free
rate). The seven predictors span the market itself, two investment *style*
factors (size and value), a *sector* factor, and three *macroeconomic* signals
(the yield-curve slope, the corporate credit spread, and market-wide volatility).

## Dataset and Variables

**Source.** All data is pulled programmatically through public market-data APIs.
Daily adjusted closing prices come from a market-price API; the macroeconomic and
interest-rate series come from the U.S. Federal Reserve's FRED API. A small R
pipeline (`R/update_data.R`) fetches every series, caches it, and assembles the
clean daily panel used here.

**What each row represents.** One **trading day**. Each row pairs Micron's return
for that day with the value of every market, style, sector, and macro factor on
the same day.

**Number of observations.** The model is estimated on **`r n_obs` trading days**,
spanning **`r date_min`** to **`r date_max`**.

**Variables.**

| Role | Variable | What it represents |
|------|----------|--------------------|
| Outcome | **MU excess return** | Micron's daily return minus the risk-free rate. |
| Predictor 1 | **Market** | The total stock market's excess return (a broad-market ETF minus the risk-free rate). This is the classic CAPM market factor. |
| Predictor 2 | **Size** | Small-cap return minus large-cap return. Positive when smaller companies outperform. |
| Predictor 3 | **Value** | Value-stock return minus growth-stock return. Positive when cheap "value" stocks beat fast-growing "growth" stocks. |
| Predictor 4 | **Sector** | The semiconductor sector's return above the market. Captures chip-industry-specific moves. |
| Predictor 5 | **TermSpread** | The daily change in the 10-year-minus-3-month Treasury yield slope. A rising slope often signals improving growth expectations. |
| Predictor 6 | **CreditSpread** | The daily change in the Baa-corporate-minus-10-year-Treasury yield spread. A rising spread signals growing financial stress. |
| Predictor 7 | **Volatility** | The daily change in the VIX, the market's "fear gauge." |

The risk-free rate is the 3-month Treasury bill rate. The spreads and the VIX
enter as **daily changes** rather than levels, so each predictor is a stationary
daily signal that lines up naturally with a daily return.

## The Regression Model Equation

The general form of a multiple linear regression with seven predictors is:

$$
Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_3 X_3 + \beta_4 X_4
      + \beta_5 X_5 + \beta_6 X_6 + \beta_7 X_7 + \varepsilon
$$

where $Y$ is the outcome, $\beta_0$ is the intercept, each $\beta_k$ is the
coefficient on predictor $X_k$, and $\varepsilon$ is the error term. Written with
our own variable names:

$$
\begin{aligned}
\text{MU excess} = {} & \beta_0 + \beta_1\,\text{Market} + \beta_2\,\text{Size}
  + \beta_3\,\text{Value} + \beta_4\,\text{Sector} \\
  & + \beta_5\,\text{TermSpread} + \beta_6\,\text{CreditSpread}
  + \beta_7\,\text{Volatility} + \varepsilon
\end{aligned}
$$

After fitting the model, the estimated equation (using the actual coefficients
computed below) is:

$$
\begin{aligned}
\widehat{\text{MU excess}} = {} & `r fmt(co[1])`
  + `r fmt(co[2])`\,\text{Market}
  `r fmt(co[3])`\,\text{Size}
  + `r fmt(co[4])`\,\text{Value}
  + `r fmt(co[5])`\,\text{Sector} \\
  & + `r fmt(co[6])`\,\text{TermSpread}
  `r fmt(co[7])`\,\text{CreditSpread}
  `r fmt(co[8])`\,\text{Volatility}
\end{aligned}
$$

## Running the Model in R

```{r run-model}
#| code-fold: false
model <- lm(MU_excess ~ Market + Size + Value + Sector +
              TermSpread + CreditSpread + Volatility, data = model_df)
summary(model)
```

```{r coef-table}
broom::tidy(model) |>
  mutate(across(where(is.numeric), \(x) round(x, 5))) |>
  kable(caption = "Estimated coefficients, standard errors, t-statistics, and p-values.")
```

## Interpreting the Model Summary

### A. Coefficients

Each coefficient is the average change in Micron's daily excess return associated
with a one-unit increase in that predictor, **holding the other six predictors
constant**.

- **Intercept (`r fmt(co["(Intercept)"])`).** In an excess-return factor model the
  intercept is the model's **alpha**: the average daily return Micron earns *beyond*
  what its factor exposures explain. Here it is tiny — about
  `r fmt(100 * co["(Intercept)"], 3)`% per day — and statistically indistinguishable
  from zero. This is the expected result: once the market, sector, and style factors
  are accounted for, Micron shows no systematic leftover drift, so the explanatory
  work is done by the factors rather than by an unexplained constant.
- **Market (`r fmt(co["Market"])`).** Positive and modestly above one. Holding the
  other predictors constant, a one-unit increase in the market's excess return is
  associated with an increase of about `r fmt(co["Market"])` in Micron's expected
  excess return. A loading above one means Micron amplifies market moves, but at
  `r fmt(co["Market"], 2)` over the full sample the amplification is slight. This is
  a sample average; the structural-break analysis below shows the market loading
  rose in the post-2024 regime.
- **Size (`r fmt(co["Size"])`).** Negative. When small-cap stocks beat large-caps,
  Micron tends to lag slightly — consistent with Micron trading like a large-cap.
- **Value (`r fmt(co["Value"])`).** Positive. When value stocks beat growth stocks,
  Micron tends to *rise*. Unlike the pure megacap growth names, Micron carries a
  modest **value** tilt — fitting for a capital-intensive, cyclical memory
  manufacturer whose earnings swing with the commodity-like price of chips rather
  than with a smooth growth narrative.
- **Sector (`r fmt(co["Sector"])`).** Strongly positive and the largest loading in
  the model. Independent of the broad market, a one-unit increase in the
  semiconductor sector's excess return is associated with a roughly one-for-one
  increase in Micron's return, indicating that the chip sector is its dominant
  systematic exposure.
- **TermSpread (`r fmt(co["TermSpread"])`).** A one-day steepening of the yield
  curve carries a small positive association with Micron's return, consistent with
  a cyclical name that benefits when the market prices in stronger growth.
- **CreditSpread (`r fmt(co["CreditSpread"])`).** Negative. On days when the
  corporate credit spread widens — a sign of rising financial stress — Micron
  tends to fall, even after accounting for the market.
- **Volatility (`r fmt(co["Volatility"])`).** Very small once the market factor is
  in the model, because changes in the VIX and market returns carry overlapping
  information.

### B. P-values

```{r pvals}
broom::tidy(model) |>
  transmute(term, estimate = round(estimate, 5), p.value = signif(p.value, 3),
            significant_5pct = ifelse(p.value < 0.05, "yes", "no")) |>
  kable(caption = "Which predictors show statistically significant relationships.")
```

A small p-value is evidence that a predictor's true coefficient is different from
zero. The **market, size, value, sector, and term-spread** factors are all
significant at the 5% level: the data give strong evidence that each is genuinely
related to Micron's return, and the **sector** and **market** loadings in
particular are estimated with overwhelming confidence. By contrast, the **credit
spread** and the **change in the VIX** are *not* statistically significant for
Micron: once the market and sector factors are present, these two add no
detectable independent relationship with Micron's daily return. A non-significant
predictor does not prove there is no relationship — only that this data set shows
no clear evidence of one.

The **intercept** is likewise not significant, which is the desired result for an
excess-return model: it indicates Micron earns no reliable daily return beyond what
its factor exposures explain — no "free lunch" left unaccounted for by the model.

### C. R-squared

The R-squared value is **`r fmt(s$r.squared, 3)`**, which means the model explains
approximately **`r round(100 * s$r.squared)`%** of the day-to-day variation in
Micron's excess return. In other words, roughly that share of Micron's daily moves
is *systematic* — traceable to broad market, style, sector, and macro forces —
while the remaining **`r round(100 * (1 - s$r.squared))`%** is **idiosyncratic**:
specific to Micron (memory-pricing news, earnings, guidance, supply decisions) and
not captured by these factors.

### D. Adjusted R-squared

The adjusted R-squared is **`r fmt(s$adj.r.squared, 3)`**. Adjusted R-squared
penalizes the model for each additional predictor, so it rises only when a new
variable adds explanatory power beyond what its degrees of freedom cost. Here it is
almost identical to the ordinary R-squared (`r fmt(s$r.squared, 3)`), indicating that
the predictor count is not inflating the fit. This is a statement about the set of
predictors as a whole, not about each one individually: with `r n_obs` observations,
one predictor costs a negligible fraction of a degree of freedom, so a flat adjusted
R-squared is consistent with two of the seven predictors — the credit spread and the
change in the VIX — being statistically insignificant. The overall fit is not an
artifact of model size, but not every predictor contributes detectable independent
explanatory power.

## Residual Errors and Model Fit

A **residual** is the difference between the actual value and the model's
prediction:

$$\text{Residual} = \text{Actual} - \text{Predicted}$$

A residual tells us how far off the model was on a single day. A **positive**
residual means the model **underpredicted** (Micron did better than expected); a
**negative** residual means it **overpredicted** (Micron did worse than expected).
For example, if Micron's actual excess return on a day was 3% and the model
predicted 2%, the residual is +1%.

## Residual Plot

```{r resid-plot}
#| fig-cap: "Residuals versus fitted values. A random horizontal band around zero indicates a well-behaved linear model."
model_df$predicted <- fitted(model)
model_df$residuals <- residuals(model)

ggplot(model_df, aes(x = predicted, y = residuals)) +
  geom_point(alpha = 0.25, color = "#2c7fb8") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(se = FALSE, color = "#d95f02", linewidth = 0.7) +
  labs(title = "Residuals vs. Predicted Values (Micron excess return)",
       x = "Predicted excess return", y = "Residual") +
  theme_minimal()
```

```{r diagnostic-plots}
#| fig-cap: "Standard regression diagnostics: residuals vs fitted, Normal Q-Q, scale-location, and residuals vs leverage."
par(mfrow = c(2, 2)); plot(model); par(mfrow = c(1, 1))
```

The residuals are scattered fairly evenly above and below zero with no obvious
trend across the range of predictions, so the orange smoother stays close to the
dashed zero line — the **linearity** assumption looks reasonable. The visible
feature is that the *spread* of the residuals widens on high-volatility days:
large surprises cluster together, a familiar property of stock returns and an
especially pronounced one for a cyclical memory name whose earnings days produce
outsized moves. The model does not systematically over- or under-predict in any
region; its errors are roughly centered on zero everywhere, but it is least
precise during turbulent market periods. The Q-Q plot shows heavier tails than a
normal distribution — extreme days (large jumps) happen more often than a bell
curve would predict.

## Model Assumptions and Limitations

```{r diagnostics}
vif_vals <- car::vif(model)
bp <- lmtest::bptest(model)
dw_lag1 <- cor(head(residuals(model), -1), tail(residuals(model), -1))
```

- **Linearity.** The residual plot shows no strong curvature, so a straight-line
  relationship between the factors and Micron's return is a defensible
  approximation.
- **Residual randomness / independence.** The day-to-day autocorrelation of the
  residuals is only about `r fmt(dw_lag1, 3)`, essentially zero, so the model is
  not leaving obvious predictable structure in its errors.
- **Outliers.** Extreme trading days (earnings surprises, market shocks) appear in
  the tails of the Q-Q plot. They do not dominate the fit given the large sample,
  but they are the days the model predicts least well.
- **Multicollinearity.** All variance-inflation factors are low (the largest is
  about `r fmt(max(vif_vals), 1)`, well under the usual threshold of 5), so the
  predictors are not redundant and each coefficient is estimated on its own
  distinct information.
- **Constant error variance.** A Breusch–Pagan test rejects constant variance
  (p `r ifelse(bp$p.value < 0.001, "< 0.001", paste("=", fmt(bp$p.value, 3)))`):
  the model's errors are larger on volatile days. This **heteroskedasticity** does
  not bias the coefficients, but it means the reported standard errors slightly
  overstate the precision of the estimates. The large significances here are wide
  enough that this caveat does not change which factors matter.
- **Causation.** This model shows **association, not causation**. A significant
  coefficient means a factor *moves together with* Micron's return, not that it
  *causes* it. These factors and Micron are all driven by the same underlying
  market forces in real time.

## Has the AI Era Changed Micron's Character?

The model above estimates a single set of factor loadings over the entire sample,
which is valid only if those loadings are stable through time. Micron is a memory
producer, and large-scale AI computing — a major new source of demand for
high-bandwidth memory (HBM) — is a plausible reason for its loadings to have
shifted. If they did, a full-sample regression reports an average across two
distinct regimes and fits each one less well than a regime-specific model would.
This section tests loading stability, locates any break date from the data,
and quantifies what changed.

### An AI-demand factor

Exposure to AI computing demand enters as a daily signal through NVIDIA's return,
which serves as a bellwether for data-center AI spending — the same spending that
drives HBM demand at Micron. Much of NVIDIA's daily move is already captured by the
market and semiconductor factors, so the raw return would be collinear with them.
The **AI** factor is therefore the residual of NVIDIA's return after regressing it
on the market and sector factors: the orthogonal component that those two factors
do not explain. This isolates the incremental AI-demand signal and keeps it
distinct from the existing predictors.

```{r ai-factor}
#| include: false
brk_df <- model_df |> arrange(date)
# Orthogonalize NVIDIA against the market and sector so AI carries only the
# incremental AI-demand signal not already in those factors.
brk_df$AI <- residuals(lm(NVDA_ret ~ Market + Sector, data = brk_df))

base_form <- MU_excess ~ Market + Size + Value + Sector +
  TermSpread + CreditSpread + Volatility
ai_form   <- update(base_form, . ~ . + AI)
```

### Detecting the break

The break date is estimated from the data rather than fixed in advance, since
choosing it by inspection would bias any test evaluated at that point. The **supF
(Quandt–Andrews) test** computes a Chow break statistic at every candidate date in
the interior of the sample, takes the maximum, and compares it against a critical
value adjusted for having searched over all dates. It tests the null of stable
coefficients against a single unknown break and, if rejected, identifies the most
likely break date.

```{r supf-test}
fs   <- strucchange::Fstats(ai_form, data = brk_df, from = 0.15)
supf <- strucchange::sctest(fs)
brk_date <- brk_df$date[fs$breakpoint]

bp_obj <- strucchange::breakpoints(ai_form, data = brk_df, h = 0.15)
n_breaks <- sum(!is.na(bp_obj$breakpoints))
```

The supF statistic is **`r fmt(as.numeric(supf$statistic), 1)`** with a p-value of
**`r ifelse(supf$p.value < 0.001, "< 0.001", fmt(supf$p.value, 3))`**, rejecting
coefficient stability at any conventional level. The estimated break date is
**`r format(brk_date)`**. A Bai–Perron procedure that allows for multiple breaks
selects **`r n_breaks`** as the optimal number, at the same date, so the evidence
points to a single regime change rather than gradual drift.

### Why that date

The estimated break falls in **March 2024**, the quarter in which Micron's guidance
shifted to HBM capacity selling out and AI data-center demand as the primary driver
of its business. This coincides with the market's re-rating of Micron from a
cyclical commodity-memory producer toward an AI-memory name. The break therefore
corresponds to a datable change in the company's business rather than a purely
statistical artifact. The chart below shows Micron's rolling factor loadings over
the sample, with relevant AI milestones marked.

```{r rolling-betas}
#| fig-cap: "Rolling 126-trading-day factor loadings for Micron. Vertical lines mark AI milestones; the dashed line is the statistically detected break."
#| fig-width: 7.5
#| fig-height: 4
win <- 126L
roll_betas <- function(i) {
  idx <- (i - win + 1):i
  cf <- coef(lm(MU_excess ~ Market + Sector + AI, data = brk_df[idx, ]))
  c(Market = cf["Market"], Sector = cf["Sector"], AI = cf["AI"])
}
roll <- t(vapply(win:nrow(brk_df), roll_betas, numeric(3)))
roll_df <- tibble::tibble(
  date   = brk_df$date[win:nrow(brk_df)],
  Market = roll[, 1], Sector = roll[, 2], AI = roll[, 3]
) |>
  tidyr::pivot_longer(-date, names_to = "Factor", values_to = "Beta")

milestones <- tibble::tibble(
  date  = as.Date(c("2022-11-30", "2024-03-20")),
  label = c("ChatGPT", "MU HBM / AI guidance")
)

ggplot(roll_df, aes(date, Beta, color = Factor)) +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_vline(xintercept = brk_date, linetype = "dashed", color = "grey40") +
  geom_vline(data = milestones, aes(xintercept = date),
             linetype = "dotted", color = "grey55") +
  geom_text(data = milestones, aes(x = date, y = Inf, label = label),
            inherit.aes = FALSE, angle = 90, vjust = -0.3, hjust = 1,
            size = 2.7, color = "grey45") +
  geom_line(linewidth = 0.7) +
  scale_color_manual(values = c(Market = "#1b9e77", Sector = "#7570b3",
                                AI = "#d95f02")) +
  labs(title = "Micron's factor loadings drifted as the AI era arrived",
       x = NULL, y = "Rolling beta (126-day window)") +
  theme_minimal() +
  theme(plot.margin = margin(t = 18, r = 8, b = 4, l = 4))
```

The rolling loadings show Micron's sensitivity to the broad market and the chip
sector rising over the sample, with the AI-demand loading increasing distinctly
around the estimated break.

### What changed, and by how much

To quantify the shift, we fit a single model over the full sample with a **Post**
indicator (one from the break date onward) interacted with the factors most likely
to have moved. Each interaction coefficient is the *change* in that loading after
the break. Because the residuals are heteroskedastic, the coefficients are tested
with heteroskedasticity-robust (HC1) standard errors.

```{r interaction-model}
brk_df$Post <- as.integer(brk_df$date >= brk_date)
inter <- lm(MU_excess ~ (Market + Sector + Value + AI) * Post +
              Size + TermSpread + CreditSpread + Volatility, data = brk_df)

robust_se <- lmtest::coeftest(inter, vcov. = sandwich::vcovHC(inter, type = "HC1"))
broom::tidy(robust_se) |>
  filter(grepl(":Post$", term) | term == "AI") |>
  mutate(across(where(is.numeric), \(x) round(x, 4))) |>
  kable(caption = "Post-break changes in Micron's factor loadings (robust standard errors). A significant interaction means the loading shifted after the AI-era break.")
```

```{r subsample-table}
periods <- lapply(c(0, 1), function(p) {
  d  <- brk_df[brk_df$Post == p, ]
  m  <- lm(base_form, data = d)
  ma <- lm(ai_form,  data = d)
  tibble::tibble(
    Period   = if (p == 0) "Pre-break" else "Post-break",
    N        = nrow(d),
    `Market beta` = round(coef(m)["Market"], 3),
    `Sector beta` = round(coef(m)["Sector"], 3),
    `AI beta`     = round(coef(ma)["AI"], 3),
    `R-squared`   = round(summary(m)$r.squared, 3)
  )
})
ai_pre  <- periods[[1]]$`AI beta`; ai_post <- periods[[2]]$`AI beta`
mkt_pre <- periods[[1]]$`Market beta`; mkt_post <- periods[[2]]$`Market beta`
sec_pre <- periods[[1]]$`Sector beta`; sec_post <- periods[[2]]$`Sector beta`
r2_pre  <- periods[[1]]$`R-squared`;   r2_post  <- periods[[2]]$`R-squared`

# Raw-return volatility and tail-frequency, by regime, to characterize the
# change in how violently Micron trades (annualized from daily returns).
mu_raw <- prices_wide |> arrange(date) |>
  transmute(date, MU = MU / dplyr::lag(MU) - 1) |>
  tidyr::drop_na() |>
  dplyr::semi_join(brk_df, by = "date") |>
  dplyr::mutate(Post = as.integer(date >= brk_date))
vol_pre   <- sd(mu_raw$MU[mu_raw$Post == 0]) * sqrt(252) * 100
vol_post  <- sd(mu_raw$MU[mu_raw$Post == 1]) * sqrt(252) * 100
big_pre   <- mean(abs(mu_raw$MU[mu_raw$Post == 0]) > 0.05) * 100
big_post  <- mean(abs(mu_raw$MU[mu_raw$Post == 1]) > 0.05) * 100

dplyr::bind_rows(periods) |>
  kable(caption = "Micron's factor loadings and fit, estimated separately before and after the detected break.")
```

Among the interacted loadings, only the **AI-demand factor shows a statistically
significant change** at the break. Micron's loading on the orthogonal AI signal
roughly doubles, from `r fmt(ai_pre, 2)` before the break to `r fmt(ai_post, 2)`
after, and the change remains significant under heteroskedasticity-robust standard
errors. The market and sector loadings also increase, but their shifts are not
individually distinguishable from zero once the AI factor is included. The regime
change is thus concentrated in Micron's sensitivity to AI computing demand rather
than spread across its broad-market or sector exposures.

The post-break R-squared is lower than the pre-break value. As Micron's returns
became more tied to AI-specific developments, a larger share of its daily variation
moved outside the span of the fixed factor set and into the idiosyncratic
component. The stock both loaded more heavily on AI demand and became harder to
explain with any constant set of factors.

The regime change is also visible in the raw scale of Micron's moves (Table 5).
Annualized return volatility rose from about **`r round(vol_pre)`%** before the break
to **`r round(vol_post)`%** after, and the share of trading days with an absolute
move larger than 5% rose from **`r fmt(big_pre, 1)`%** to **`r fmt(big_post, 1)`%** —
roughly one day in six in the current regime. Micron did not merely re-rate to a
higher loading on AI demand; it became a materially more volatile instrument, with
double-digit single-day moves now a routine feature of how the market prices its
exposure to AI capacity.

```{r volatility-table}
tibble::tibble(
  Period = c("Pre-break", "Post-break"),
  `Annualized volatility` = sprintf("%.0f%%", c(vol_pre, vol_post)),
  `Days with >5% move`    = sprintf("%.1f%%", c(big_pre, big_post))
) |>
  kable(caption = "Raw-return volatility and large-move frequency, before and after the detected break.")
```

Taken together, the results give a specific characterization of the regime change:
Micron's factor loadings were not stable across the sample, the instability is
dated to March 2024 and aligns with a concrete shift in the company's business, and
its principal manifestation is an approximately doubled sensitivity to AI computing
demand. Treating the change this way — as a measurable shift in existing loadings
plus one orthogonal demand factor — keeps each component identified and avoids
attributing the move to a single undated event.

## The "So What?"

The full-sample model answers the question it was built to answer — over the whole
period, the **semiconductor sector** is Micron's dominant exposure and its market
beta sits modestly above one — but those coefficients average across two distinct
regimes, and the structural-break analysis shows the average understates what
Micron has become. The relevant description of the stock today is the post-2024
regime, not the blended average.

In the current regime, Micron is a substantially more aggressive instrument. Its
market beta is about **`r fmt(mkt_post, 2)`** (versus `r fmt(mkt_pre, 2)` before the
break), its semiconductor-sector loading is **`r fmt(sec_post, 2)`**, and its loading
on the orthogonal AI-demand factor has roughly doubled to **`r fmt(ai_post, 2)`**.
The market now prices Micron heavily as a proxy for AI-memory buildout: it amplifies
both the broad market and the chip sector, and it responds directly to swings in
AI-demand sentiment. Notably, the systematic share of its variation *fell* even as
these loadings rose — the model's R-squared drops from **`r fmt(r2_pre, 2)`** before
the break to **`r fmt(r2_post, 2)`** after — so Micron became both more AI-levered
and more idiosyncratic, with a larger fraction of its daily moves now driven by
AI- and company-specific news no fixed factor set can capture. That shift comes with
far larger moves — annualized volatility near **`r round(vol_post)`%** and an
absolute daily move above 5% on roughly **`r fmt(big_post, 1)`%** of trading days,
including frequent double-digit single-session swings.

For an investor, the practical implications follow from the current regime, not the
historical average. Micron today is close to a leveraged, high-volatility bet on the
semiconductor sector and the AI-memory cycle: an investor holding the chip sector
already carries much of its systematic risk, but in amplified form, and with sizable
AI-sentiment exposure on top. Its risk is no longer that of a "mildly high-beta"
name — it is a stock that routinely moves several percent in a day and is repriced
sharply on AI-demand news. Three caveats bound the analysis: it measures
association, not causation; its errors are largest on the most volatile days; and,
most importantly here, its loadings are regime-dependent — the March 2024 break is
direct evidence that they can shift again as the AI cycle evolves. Like any factor
model, it explains the *systematic* part of Micron's returns rather than predicting
any single day's move, since a large idiosyncratic share remains by construction.
Within that scope it does its job: it identifies which forces drive Micron,
quantifies how strongly each one does, shows that those relationships shifted in
March 2024, and delivers a clear, current picture of what the stock is exposed to
and how violently it now trades.
